The hydra and the rooster

### 2020-07-08 latex/chapter-skippable.tex
add Omega_ordinal

\subsection{Notation systems for some small ordinals}

@@

\subsubsection{Finite ordinals}
Let $n$ be some natural number. The segment associated with $n$ is the interval 
$[0,n[\,=\,\{0,1,\dots,n-1\}$.

@@

\begin{remark}
 There is no interesting arihmetic on finite ordinals, since functions like successor, addition;, etc.,  cannot represented in \coq{} as \emph{total} functions.
\end{remark}

\begin{itemize}
\item $i\;(i\in\mathbb{N})$ \coq's type \verb@{j:nat | j < i}@ (see also
\href{https://www.math.nagoya-u.ac.jp/~garrigue/lecture/2018_AW/mathcomp-1.7.0/htmldoc/mathcomp.ssreflect.fintype.html#Ordinal}{finite ordinals in Mathcomp}).
\item $\omega$: \coq's type \texttt{nat}, with the order \texttt{Nat.lt}.
\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap/T1}).
\item $\Gamma_0$: The set of terms in Veblen normal form.
\end{itemize}

>>>>

\section{Notation systems for some small ordinals}

@@

\subsection{Finite ordinals}
Let $n$ be some natural number. The segment associated with $n$ is the interval 
$[0,n[\,=\,\{0,1,\dots,n-1\}$.

@@

\begin{remark}
 There is no interesting arihmetic on finite ordinals, since functions like successor, addition, etc.,  cannot be represented in \coq{} as \emph{total} functions.
\end{remark}

\paragraph{Related work}
Finite ordinals are also formailzed in MathComp~\cite{SSR}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.

\subsection{The first infinite ordinal : \texorpdfstring{$\omega$}{omega}}

Mathematically speaking, the ordinal $\omega$ is (set-theoretically) considered as the well-ordered set of
natural numbers. 
Every beginner in \coq{} must have seen examples using the type \texttt{nat} of natural numbers. \coq's standard 
library already contains a lot of functions and lemmas on Peano arithmetic.

In order to build a notation system for $\omega$, we have just to look for lemmas about strict orders and well-foundedness in type \texttt{nat}.

\begin{Coqsrc}
Require Import Arith Compare_dec Lia Simple_LexProd Ordinal_generic.
Import Relations.

Search (@StrictOrder nat).
Search @Nat.compare.
Search (@well_founded nat).
\end{Coqsrc}

\begin{Coqsrc}
Instance Omega : OrdinalNotation Nat.lt_strorder Nat.compare.
Proof.
 split.
 - apply Nat.compare_spec.
 - apply lt_wf.
Qed.
\end{Coqsrc}

\begin{itemize}
\item $\omega$: \coq's type \texttt{nat}, with the order \texttt{Nat.lt}.
\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap/T1}).
\item $\Gamma_0$: The set of terms in Veblen normal form.
\end{itemize}
